---
title: "wk6prac"
output: html_document
date: "2022-11-16"
---

# Key Advice

Select points or polygons in a polygon = Selecting data by location = spatial sub-setting

Determine where datasets overlap (or touch, or don’t overlap) and extract those parts = spatial clipping

Join two spatial datasets together = spatial joining, which can use spatial subsetting functions as the default is st_intersects(). This function joins spatial data.

Select data by attributes = filtering or selecting rows / columns with dplyr

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r libraries}
library(spatstat)
library(here)
library(sp)
library(rgeos)
library(maptools)
library(tmap)
library(sf)
library(geojson)
library(geojsonio)
library(tmaptools)
library(tidyverse)
library(dplyr)
library(stringr)
library(sf)
```

# Read in Data

```{r pressure, echo=FALSE}

LondonBoroughs <- st_read("https://opendata.arcgis.com/datasets/8edafbe3276d4b56aec60991cbddda50_4.geojson") %>% 
  st_transform(., 27700)
  
BluePlaques <- st_read("https://s3.eu-west-2.amazonaws.com/openplaques/open-plaques-london-2018-04-08.geojson") %>% 
  st_transform(., 27700)

OSM <- st_read(here::here("prac6_data", "greater-london-latest-free.shp",   "gis_osm_pois_a_free_1.shp")) %>%
  st_transform(., 27700) %>%
  filter(fclass == 'hotel')

Londonborough <- st_read("/Users/martignoni/Documents/CASA/GIS/wk1/statistical-gis-boundaries-london/ESRI/London_Borough_Excluding_MHW.shp") %>% 
  st_transform(.,27700)

```

# Setting Up Data

```{r}

#Filtering Data
BoroughMap <- LondonBoroughs %>% 
  dplyr::filter(str_detect(lad15cd, "^E09"))

#Checking the Maps
qtm(BoroughMap)

tmap_mode("plot")
tm_shape(BoroughMap) + tm_polygons(col = NA, alpha = 0.5) + tm_shape(BluePlaques) + tm_dots(col = "blue")

#Cleaning Data
BluePlaques <-  distinct(BluePlaques)

#Spatial Subsetting
BluePlaquesSub <- BluePlaques[BoroughMap,]
#This tells R to select for the points located within the BoroughMap geometry

tmap_mode("plot")
tm_shape(BoroughMap)+tm_polygons(col=NA, alpha=0.5)+tm_shape(BluePlaquesSub)+tm_dots(col="blue")

intersect_indices <-st_intersects(BoroughMap, BluePlaques)
  
```

# Spatial Clipping

```{r}

Londonborough <- st_read("/Users/martignoni/Documents/CASA/GIS/wk1/statistical-gis-boundaries-london/ESRI/London_Borough_Excluding_MHW.shp") %>% 
  st_transform(.,27700)

```

# Spatial Joining

```{r}
#Do an example of a spatial join (the OSM and landmark(Hotels) data)
join_example <- st_join(OSM, Londonborough)
nrow(join_example)

# Now, we're going to read in the .csv BUT the data is weird so we gotta make the coordinate system into lon/lat and therefore make it inot spatial data

Airbnb <- read_csv("/Users/martignoni/Documents/CASA/GIS/wk5/wk5prac/wk5_data/listings.csv") %>%
  st_as_sf(., coords = c("longitude", "latitude"),crs = 4326) %>% 
  st_transform(., 27700)%>%
  filter(room_type == 'Entire home/apt' & availability_365 =='365')


# A function for the join functions are covered in practical 7

Joinfun <- function(data1, data2){
output<- data1%>%
  st_join(Londonborough,.) %>%
  add_count(GSS_CODE, name="hotels_in_borough") 
  return(output)}

# function for hotels
Hotels <- Joinfun(OSM, Londonborough)

# function for airbnb
Airbnb <- Joinfun(Airbnb, Londonborough)
 
Hotels <- Hotels %>%
  #at the moment each hotel is a row for the borough
  #we just one one row that has number of airbnbs
  group_by(., GSS_CODE, NAME)%>%
  summarise('Accomodation count' = unique(hotels_in_borough))

Airbnb <- Airbnb %>%
  group_by(., GSS_CODE, NAME)%>%
  summarise('Accomodation count' = unique(hotels_in_borough))

all_accomodation <- st_join(Hotels, Airbnb)
head(all_accomodation) #the st_join does not quite work here because name.x and name.y don't match. st_join is a left join, here the left table is hotels. SO, use st_equals

all_accomodation1 <- st_equals(Hotels, Airbnb)
```

# Study Area - zooming in

```{r}
#Select specifically for the Borough of Harrow because it will be easier to focus on one of them
Harrow <- BoroughMap %>%
  filter(., lad15nm=="Harrow")

#check to see you've selected harrow
tm_shape(Harrow) + tm_polygons(col = NA, alpha = 0.5)

#clip the blue plaques data
BluePlaquesSub <- BluePlaques[Harrow,]
tmap_mode("plot")
tm_shape(Harrow)+tm_polygons(col=NA, alpha=0.5)+tm_shape(BluePlaquesSub)+tm_dots(col="blue")

#Now, we now want to look at spatial stats (spatstat package) and to do that we need a window
window <- as.owin(Harrow)
plot(window)

#spatstat doesn't work with sp or sf, it uses its own datatype: ppp (point pattern object). First we make a sp object, then a ppp.
BluePlaquesSub <- BluePlaquesSub %>% 
  as(., 'Spatial')

BluePlaquesSub.ppp <- ppp(x=BluePlaquesSub@coords[,1], y=BluePlaquesSub@coords[,2], window=window) #makes the ppp

#Now to visualise the ppp
BluePlaquesSub.ppp %>% 
  plot(.,pch=16,cex=0.5, main="Blue Plaques Harrow")

```

# Point Pattern Analysis
## Kernel Density Estimation (KDE)
```{r}
#This is a Kernel Density Estimation (KDE) map; sigma value sets the diameter of the Kernel
BluePlaquesSub.ppp %>% 
  density(., sigma=500) %>% 
  plot()
```

## Quadrat Analysis

```{r}
#Quadrat Analysis

#First, plot points:
plot(BluePlaquesSub.ppp, pch=16,cex=0.5, main="Blue Plaques in Harrow")

#Second, count the points in that fall in a 6 x 6 grid overlaid across the window
BluePlaquesSub.ppp %>% 
  quadratcount(., nx=6, ny=6) %>% 
  plot(., add=T, col="red")

## ASK ABOUT FROM HERE BELOW ##

#quadratcount can also give us a table of the counts
Qcount <- BluePlaquesSub.ppp %>%
  quadratcount(.,nx = 6, ny = 6) %>%
  as.data.frame() %>%
  dplyr::count(Var1=Freq)%>%
  dplyr::rename(Freqquadratcount=n)

Qcount %>% 
  summarise_all(class)

#Now we have a table and we need to check it against the possion distribution

sums <- Qcount %>% #this is the total number of plaques
  mutate(total=Var1*Freqquadratcount) %>% 
  dplyr::summarise(across(everything(),sum)) %>% 
  dplyr::select(-Var1)

lambda <- Qcount %>% #here we are calculating the lambda for the Poisson Dist
  mutate(total=Var1*Freqquadratcount) %>% 
  dplyr::summarise(across(everything(), sum)) %>%
  mutate(lambda=total/Freqquadratcount) %>% 
  dplyr::select(lambda) %>% 
  pull(lambda)

QcountTable <- Qcount %>% #here we are calculating the expected count based on the total number of plaques
  mutate(Pr=((lambda^Var1)*exp(-lambda))/factorial(Var1))%>% 
  mutate(Expected= (round(Pr * sums$Freqquadratcount, 0)))
  
#plot!
plot(c(1,5),c(0,14), type="n", xlab="Number of Blue Plaques (Red=Observed,Blue=Expected)", ylab="Frequency of Occurances")
points(QcountTable$Freqquadratcount, col="Red", type="o",  lwd=3)
points(QcountTable$Expected, col="Blue", type="o", lwd=3)

```

## Ripley’s K

```{r}
K <- BluePlaquesSub.ppp %>% 
  Kest(., correction="border") %>% 
  plot()
```


# Density-based spatial clustering of applications with noise: DBSCAN

```{r}

```


