---
title: "wk6prac"
output: html_document
date: "2022-11-16"
---

# Key Advice

Select points or polygons in a polygon = Selecting data by location = spatial sub-setting

Determine where datasets overlap (or touch, or don’t overlap) and extract those parts = spatial clipping

Join two spatial datasets together = spatial joining, which can use spatial subsetting functions as the default is st_intersects(). This function joins spatial data.

Select data by attributes = filtering or selecting rows / columns with dplyr

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r libraries}
library(spatstat)
library(here)
library(sp)
library(rgeos)
library(maptools)
library(tmap)
library(sf)
library(geojson)
library(geojsonio)
library(tmaptools)
library(tidyverse)
library(dplyr)
library(stringr)
library(sf)
```

# Read in Data

```{r pressure, echo=FALSE}

LondonBoroughs <- st_read("https://opendata.arcgis.com/datasets/8edafbe3276d4b56aec60991cbddda50_4.geojson") %>% 
  st_transform(., 27700)
  
BluePlaques <- st_read("https://s3.eu-west-2.amazonaws.com/openplaques/open-plaques-london-2018-04-08.geojson") %>% 
  st_transform(., 27700)

OSM <- st_read(here::here("prac6_data", "greater-london-latest-free.shp",   "gis_osm_pois_a_free_1.shp")) %>%
  st_transform(., 27700) %>%
  filter(fclass == 'hotel')

Londonborough <- st_read("/Users/martignoni/Documents/CASA/GIS/wk1/statistical-gis-boundaries-london/ESRI/London_Borough_Excluding_MHW.shp") %>% 
  st_transform(.,27700)

```

# Setting Up Data

```{r}

#Filtering Data
BoroughMap <- LondonBoroughs %>% 
  dplyr::filter(str_detect(lad15cd, "^E09"))

#Checking the Maps
qtm(BoroughMap)

tmap_mode("plot")
tm_shape(BoroughMap) + tm_polygons(col = NA, alpha = 0.5) + tm_shape(BluePlaques) + tm_dots(col = "blue")

#Cleaning Data
BluePlaques <-  distinct(BluePlaques)

#Spatial Subsetting
BluePlaquesSub <- BluePlaques[BoroughMap,]
#This tells R to select for the points located within the BoroughMap geometry

tmap_mode("plot")
tm_shape(BoroughMap)+tm_polygons(col=NA, alpha=0.5)+tm_shape(BluePlaquesSub)+tm_dots(col="blue")

intersect_indices <-st_intersects(BoroughMap, BluePlaques)
  
```

# Spatial Clipping

```{r}

Londonborough <- st_read("/Users/martignoni/Documents/CASA/GIS/wk1/statistical-gis-boundaries-london/ESRI/London_Borough_Excluding_MHW.shp") %>% 
  st_transform(.,27700)

```

# Spatial Joining

```{r}
#Do an example of a spatial join (the OSM and landmark(Hotels) data)
join_example <- st_join(OSM, Londonborough)
nrow(join_example)

# Now, we're going to read in the .csv BUT the data is weird so we gotta make the coordinate system into lon/lat and therefore make it inot spatial data

Airbnb <- read_csv("/Users/martignoni/Documents/CASA/GIS/wk5/wk5prac/wk5_data/listings.csv") %>%
  st_as_sf(., coords = c("longitude", "latitude"),crs = 4326) %>% 
  st_transform(., 27700)%>%
  filter(room_type == 'Entire home/apt' & availability_365 =='365')


# A function for the join functions are covered in practical 7

Joinfun <- function(data1, data2){
output<- data1%>%
  st_join(Londonborough,.) %>%
  add_count(GSS_CODE, name="hotels_in_borough") 
  return(output)}

# function for hotels
Hotels <- Joinfun(OSM, Londonborough)

# function for airbnb
Airbnb <- Joinfun(Airbnb, Londonborough)
 
Hotels <- Hotels %>%
  #at the moment each hotel is a row for the borough
  #we just one one row that has number of airbnbs
  group_by(., GSS_CODE, NAME)%>%
  summarise('Accomodation count' = unique(hotels_in_borough))

Airbnb <- Airbnb %>%
  group_by(., GSS_CODE, NAME)%>%
  summarise('Accomodation count' = unique(hotels_in_borough))

all_accomodation <- st_join(Hotels, Airbnb)
head(all_accomodation) #the st_join does not quite work here because name.x and name.y don't match. st_join is a left join, here the left table is hotels. SO, use st_equals

all_accomodation1 <- st_equals(Hotels, Airbnb)
```

# Study Area - zooming in

```{r}
#Select specifically for the Borough of Harrow because it will be easier to focus on one of them
Harrow <- BoroughMap %>%
  filter(., lad15nm=="Harrow")

#check to see you've selected harrow
tm_shape(Harrow) + tm_polygons(col = NA, alpha = 0.5)

#clip the blue plaques data
BluePlaquesSub <- BluePlaques[Harrow,]
tmap_mode("plot")
tm_shape(Harrow)+tm_polygons(col=NA, alpha=0.5)+tm_shape(BluePlaquesSub)+tm_dots(col="blue")

#Now, we now want to look at spatial stats (spatstat package) and to do that we need a window
window <- as.owin(Harrow)
plot(window)

#spatstat doesn't work with sp or sf, it uses its own datatype: ppp (point pattern object). First we make a sp object, then a ppp.
BluePlaquesSub <- BluePlaquesSub %>% 
  as(., 'Spatial')

BluePlaquesSub.ppp <- ppp(x=BluePlaquesSub@coords[,1], y=BluePlaquesSub@coords[,2], window=window) #makes the ppp

#Now to visualise the ppp
BluePlaquesSub.ppp %>% 
  plot(.,pch=16,cex=0.5, main="Blue Plaques Harrow")

```

# Point Pattern Analysis
## Kernel Density Estimation (KDE)
```{r}
#This is a Kernel Density Estimation (KDE) map; sigma value sets the diameter of the Kernel
BluePlaquesSub.ppp %>% 
  density(., sigma=500) %>% 
  plot()
```

## Quadrat Analysis

```{r}
#Quadrat Analysis

#First, plot points:
plot(BluePlaquesSub.ppp, pch=16,cex=0.5, main="Blue Plaques in Harrow")

#Second, count the points in that fall in a 6 x 6 grid overlaid across the window
BluePlaquesSub.ppp %>% 
  quadratcount(., nx=6, ny=6) %>% 
  plot(., add=T, col="red")

## ASK ABOUT FROM HERE BELOW ##

#quadratcount can also give us a table of the counts
Qcount <- BluePlaquesSub.ppp %>%
  quadratcount(.,nx = 6, ny = 6) %>%
  as.data.frame() %>%
  dplyr::count(Var1=Freq)%>%
  dplyr::rename(Freqquadratcount=n)

Qcount %>% 
  summarise_all(class)

#Now we have a table and we need to check it against the possion distribution

sums <- Qcount %>% #this is the total number of plaques
  mutate(total=Var1*Freqquadratcount) %>% 
  dplyr::summarise(across(everything(),sum)) %>% 
  dplyr::select(-Var1)
# -argument removes the column from the dataframe

lambda <- Qcount %>% #here we are calculating the lambda for the Poisson Dist
  mutate(total=Var1*Freqquadratcount) %>% 
  dplyr::summarise(across(everything(), sum)) %>%
  mutate(lambda=total/Freqquadratcount) %>% 
  dplyr::select(lambda) %>% 
  pull(lambda)

QcountTable <- Qcount %>% #here we are calculating the expected count based on the total number of plaques
  mutate(Pr=((lambda^Var1)*exp(-lambda))/factorial(Var1))%>% 
  mutate(Expected= (round(Pr * sums$Freqquadratcount, 0)))
  
#plot!
plot(c(1,5),c(0,14), type="n", xlab="Number of Blue Plaques (Red=Observed,Blue=Expected)", ylab="Frequency of Occurances")
points(QcountTable$Freqquadratcount, col="Red", type="o",  lwd=3)
points(QcountTable$Expected, col="Blue", type="o", lwd=3)

```

## Ripley’s K

```{r}
K <- BluePlaquesSub.ppp %>% 
  Kest(., correction="border") %>% 
  plot()

Kval <- as.data.frame(Kest(BluePlaquesSub.ppp, correction = "Ripley"))

```


# Density-based spatial clustering of applications with noise: DBSCAN

Notes:
1. Ripley's K and Quadrat Analysis tell us that there is clustering but not *where* the clustering is, so we do DBSCAN (or Optics or HBSCAN)

2. You can use the r found from Ripley's K or Quadrat Analysis for DBSCAN. The biggest bulge in the graph above the poisson dist is at around 700m, so we'll use that.

3. DBSCAN has two parametres: (1) Epsilon - this is the radius within which the algorithm with search for clusters (2) MinPts - this is the minimum number of points that should be considered a cluster

4. Here are the steps in writing the code:
  (1) extract the points from the spatial points data frame
  (2) run the dbscan analysis
  (3) plot the results
  
5. We can also use a Knee plot to determine the best epsilon value, that uses kNNdistplot() from dbscan. This plot shows the avg distance for points from "k" neighbours in ascending order, this is where K (dist from neighbours) increases. In choosing the knee, you can do the biggest knee or then use it to see what gives the most reasonable clusters. You can also look at the knees to help read ripley's k and what is says for r.

6. We can also use ggplot2 to make a much nicer DBSCAN. The steps for this are outlined below.


```{r}
library(raster)
library(fpc)
library(dbscan)
library(ggplot2)
library(OpenStreetMap)

st_geometry(BoroughMap) # to check the crs of the polygon

#DBSCAN step 1 (using dbscan package)
BluePlaquesSubPoints <- BluePlaquesSub %>% 
  coordinates(.) %>% 
  as.data.frame()

#DBSCAN step 2 (using dbscan package)
db <- BluePlaquesSubPoints %>% 
  fpc::dbscan(., eps = 700, MinPts = 4)

#DBSCAN step 3 (using dbscan package)
plot(db, BluePlaquesSubPoints, main="DBSCAN Output", frame = F)
plot(BoroughMap$geometry, add = T)

#Knee plot

BluePlaquesSubPoints %>% 
  dbscan::kNNdistplot(., k=4)

#Making the DBSCAN using data from Knee Plot

db1 <- BluePlaquesSubPoints %>%
  fpc::dbscan(.,eps = 500, MinPts = 4)

plot(db1, BluePlaquesSubPoints, main = "DBSCAN Output", frame = F)
plot(BoroughMap$geometry, add=T)

#Making DBSCAN using ggplot. Here, if we click on the db object, we can see a "cluster" row that we then want to add back to out dataframe.

#DBSCAN step 1 (using ggplot2 package)
BluePlaquesSubPoints<- BluePlaquesSubPoints %>%
  mutate(dbcluster=db$cluster)

#DBSCAN step 2 (using ggplot2 package), here we can make convex hull polygons around the clusters.
chulls <- BluePlaquesSubPoints %>% 
  group_by(dbcluster) %>% 
  dplyr::mutate(hull = 1:n(), hull = factor(hull, chull(coords.x1, coords.x1))) %>%
  arrange(hull)

#DBSCAN step 3 (using ggplot2 package). we need to remove zero from the clusters (in the df) because it just represents what is *not* a cluster
chulls <- chulls %>% 
  filter(dbcluster>=1)

#DBSCAN step 4 (using ggplot2 package). Now we make the plot.
dbplot <- ggplot(data=BluePlaquesSubPoints, aes(coords.x1,coords.x2, colour=dbcluster, fill=dbcluster)) 
  #add the points in
dbplot <- dbplot + geom_point()
  #now the convex hulls
dbplot <- dbplot + geom_polygon(data = chulls,aes(coords.x1,coords.x2, group=dbcluster),alpha = 0.5) 
  #now plot, setting the coordinates to scale correctly and as a black and white plot 
dbplot + theme_bw() + coord_equal()

#DBSCAN step 4 (using ggplot2 package/open stree map). Set a bounding box (in WGS84),convert back to BNG, to add a base map.
HarrowWGSbb <- Harrow %>% 
  st_transform(., 4326) %>% 
  st_bbox()

basemap <- OpenStreetMap::openmap(c(51.5549876,-0.4040502),c(51.6405356,-0.2671315),
                         zoom=NULL,
                         "stamen-toner") 
# convert BACK to bng
basemap_bng <- openproj(basemap, projection="+init=epsg:27700")

  #now the actual plotting!
autoplot.OpenStreetMap(basemap_bng)+geom_point(data=BluePlaquesSubPoints, aes(coords.x1,coords.x2, color=dbcluster, fill=dbcluster))+geom_polygon(data=chulls, aes(coords.x1,coords.x2, group=dbcluster, fill=dbcluster), alpha=0.5)

```


